---
title: "Untitled"
author: "Zachary M. Smith"
date: "4/24/2019"
output: html_document
---

# Station Locations

Import he Wallkill NHD clip created by Charles Stoll as a SpatialLinesDataFrame.
```{r}
nhd.gdb <- file.path(project.path, "data", "gis", "Wallkill_Clip_NHD.gdb")

streams.polyline <- rgdal::readOGR(dsn = nhd.gdb, layer = "WALK_0_7_NHD_CLIP")
```

Plot the Wallkill locations on an interactive map.
```{r}
leaflet(sites.df) %>% 
  addTiles() %>% 
  addPolylines(data = streams.polyline,
               weight = 2,
               opacity = 1) %>% 
  addMarkers(~longitude, ~latitude,
             popup = paste(
               paste("<b>Site ID:</b>", sites.df$site_id),
               paste("<b>Stream:</b>", sites.df$stream),
               paste("<b>Longitude:</b>", sites.df$longitude),
               paste("<b>Latitude:</b>", sites.df$latitude),
               paste("<b>Type:</b>", sites.df$type),
               sep = "<br/>"
             ))
```

## Data Summary

### Chemistry

#### 2017 RAS

```{r, echo=FALSE}
chem17.df <- chem.df %>% 
  filter(project_name == "wallkill ras 2017")

date.min.max <- paste(min(as.Date(chem17.df$date_time)), max(as.Date(chem17.df$date_time)), sep= " to ")
date.length <- max(as.Date(chem17.df$date_time)) - min(as.Date(chem17.df$date_time))
```

__Description:__
__Collected By:__ NYSDEC
__Intent of Collection:__
__Date Range:__ `r paste(min(as.Date(chem17.df$date_time)), max(as.Date(chem17.df$date_time)), sep= " to ")`
__Parameters Collected:__ `r knitr::kable(data.frame(Parameter = unique(chem17.df$chemical_name)))`


#### 2018 RAS

#### 2018 PEER
